Evaluation of LOOCV predictive abilities

### 2016-09-19 reports/agronomic_data.Rmd
Fix bug in class of resulting data.table

# Explore phenotypic data
Load packages
```{r packages, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "caret",
               "knitr")
```

@@

## All hybrids
Prepare agronomic data.
```{r Prepare_agro}
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
```

Evaluate the skewness of agronomic traits.
```{r Full_Agro_Skewness}
all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2)) %>%
  arrange(abs(Skewness)) %>%
  kable(padding = 0)
```

Evaluate the distribution of agronomic traits via kernel density plots.
```{r Full_Agro_Distribution}
all_pheno %>%
  ggplot(aes(x = Observed)) +
  geom_density() +
  facet_wrap(~ Trait, scales = "free") +
  theme_hc() +
  ylab("Density")
```

## Hybrids of parents with genomic and transcriptomic data 
bla bla

>>>>

# Explore phenotypic data
Load packages
```{r packages, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "caret",
               "knitr", "tibble")
```

@@

Prepare agronomic data.
```{r Prepare_agro}
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))

# Select hybrids for which both parents have genomic as well as transcriptomic
# records.
dent <- genos$Dent$mrna
flint <- genos$Flint$mrna
hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid

# All hybrids
all_pheno <- all_pheno %>%
  mutate(Reduced_Data = "no")

# Only hybrids for which parents have both, genomic and transcriptomic data
reduced_pheno <- all_pheno %>%
  mutate(Reduced_Data = ifelse(G %in% hybrid, yes = "yes", no = "no")) %>%
  filter(Reduced_Data == "yes")
```

## Skewness
A general **rule of thumb** to consider is that skewed data whose ratio of the 
highest value to the lowest value is greater than 20 have significant skewness.
The formula for the skewness statistic is
$$
\text{skewness}  = \frac{\sum (x_i - \bar{x})^{3}}{(n - 1)v^{3/2}},
$$
where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ is the predictor
variable, $n$ is the number of values and $\bar{x}$ is the sample mean of the 
predictor.
```{r Agro_Skewness}
all_skewness <- all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
reduced_skewness <- reduced_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
full_join(x = all_skewness, y = reduced_skewness,
          by = "Trait",
          suffix = c("_All", "_Reduced")) %>%
  kable(padding = 0)
```

## Distribution
```{r Full_Agro_Distribution, message = FALSE}
ggplot(all_pheno, aes(x = Observed)) +
  geom_histogram(fill = "grey", alpha = 0.5) +
  geom_histogram(data = reduced_pheno, color = "black") +
  facet_wrap(~ Trait, scales = "free") +
  theme_bw()
```

## Correlation
```{r Full_Agro_Correlation}
par(mar = c(0, 2, 0, 0),
    mfrow = c(1, 2))
all_pheno %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     type = "lower",
                     order = "hclust",
                     tl.srt = 45,
                     addCoef.col = "black",
                     tl.col = "black",
                     title = "All hybrids",
                     mar = c(0, 1, 1, 0))
reduced_pheno %>%
  filter(Reduced_Data == "yes") %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     type = "lower",
                     order = "hclust",
                     tl.srt = 45,
                     addCoef.col = "black",
                     tl.col = "black",
                     title = "Reduced hybrids",
                     mar = c(0, 1, 1, 0))
```
